---
output: github_document
title: "Carolina Kernels - Final Project"
---

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

```{r}
library(carolinaKernels)
# carolinaKernels::run_app()
```

The Carolina Kernels are proud to introduce an exploratory analysis tool built using RShiny to investigate special teams NFL data.

Every Sunday the NFL, through its Next Gen Stats (NGS) initiative, collects mountains of player data. Each year, the organization releases portions of this archive as part of an annual analytics competition coined the "Big Data Bowl".

This year participants were asked to investigate special teams play. For those unfamiliar with North American football, four distinct play types compose special teams: kickoffs, punts, extra points, and field goals. Kickoffs and punts transition possession from one team to another, while extra points and field goals involve scoring points by kicking the ball thru a goal (post).

Each team has different sets of players involved in special teams functions which are different than regular offensive and defensive rosters. Included in the NFL's data dump was scouting information, player movement for all special teams plays between 2018-2020, and metadata corresponding to each play (time of game, outcome, team, etc.).

Our final application allows users to explore various summary statistics for each of the play types and view actual animations for each distinct play outcome built from the player movement file. Download and install to begin exploring our product.

For detailed documentation, install the package and run:

```{r}
vignette("carolinaKernels")
```
