---
title: "AnimateGraph"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AnimateGraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(ggplot2)
library(gganimate)
library(tidyverse)
library(dplyr)
library(cowplot)
library(ggridges)
library(repr)
library(here)

options(warn = -1)
options(repr.plot.width = 15, repr.plot.height = 10)
```

```{r, read-in-data, include=FALSE}
##
# plays = read.csv(file = paste(here(),"/data/plays.csv",sep=''), header = TRUE, stringsAsFactors = FALSE)
# tracking = read.csv(file = paste(here(),"/data/tracking2020.csv",sep=''), header = TRUE, stringsAsFactors = FALSE)
plays <- readRDS("plays.rds")
tracking <- readRDS("tracking.rds")
```


Grab one play for plot testing:
```{r}
sample_gameId <- "2020092012"
sample_playId <- "1357"
sample_play <- tracking %>% filter(gameId == sample_gameId & playId == sample_playId)

# add a group (is_football to denote if tracking )
sample_play <- sample_play %>%
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))

## sample animation plot
sample_play %>%
  filter(frameId == 1) %>%
  ggplot() +
  theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), panel.background = element_rect(fill = "#c2efff")) +
  xlim(0, 120) +
  ylim(0, 53.33) +

  ## NFL Field Specific Lines
  # add yard lines every 10 yds
  annotate("segment", x = seq(20, 105, by = 10), y = 0, xend = seq(20, 105, by = 10), yend = 53.33, col = "black") +

  # set mid-field and endzone lines
  annotate("segment", x = c(60), y = 0, xend = c(60), yend = 53.33, lwd = 1.3, col = "white") +
  annotate("segment", x = c(0, 10, 110, 120), y = 0, xend = c(0, 10, 110, 120), yend = 53.33, lwd = 1.1, col = "black") +

  # add lines at locations of hash marks
  annotate("segment", x = 10, y = (29.24667 + 30.24667) / 2, xend = 110, yend = (29.24667 + 30.24667) / 2, color = "white", lwd = 1.05) +
  annotate("segment", x = 10, y = (24.08333 + 23.08333) / 2, xend = 110, yend = (24.08333 + 23.08333) / 2, color = "white", lwd = 1.05) +

  # sideline
  annotate("segment", x = 0, y = c(0, 53.33), xend = 120, yend = c(0, 53.33), lwd = 1.3, col = "black") +
  # denote 50yd, 30yd, 10yd line with text
  annotate("text", label = 50, x = 60, y = 10, colour = "black", size = 5) +
  annotate("text", label = 30, x = 80, y = 6, colour = "black", size = 5) +
  annotate("text", label = 10, x = 100, y = 2, colour = "black", size = 5) +
  annotate("text", label = 30, x = 40, y = 6, colour = "black", size = 5) +
  annotate("text", label = 10, x = 20, y = 2, colour = "black", size = 5) +
  # above labels
  annotate("text", label = 50, x = 60, y = 44, colour = "black", size = 5) +
  annotate("text", label = 30, x = 80, y = 48, colour = "black", size = 5) +
  annotate("text", label = 10, x = 100, y = 51.5, colour = "black", size = 5) +
  annotate("text", label = 30, x = 40, y = 48, colour = "black", size = 5) +
  annotate("text", label = 10, x = 20, y = 51.5, colour = "black", size = 5) +

  # field goal posts
  annotate(geom = "segment", x = 120, xend = 120, y = 160 / 6 + 18.5 / 3 / 2, yend = 160 / 6 - 18.5 / 3 / 2, color = "yellow", lwd = 1.3) +
  annotate(geom = "segment", x = 0, xend = 0, y = 160 / 6 + 18.5 / 3 / 2, yend = 160 / 6 - 18.5 / 3 / 2, color = "yellow", lwd = 1.3) +

  ## add the points last so they sit ontop of the field markings
  scale_color_manual(values = c("#ff1212", "#8f6545", "#057d11")) +
  geom_point(aes(x = x, y = y, color = team, shape = is_football), size = 3)
```

Create Test Animation...using a series of .png files use R libary to string together an animation.
We will serve this on the webpage. Could also have option for user to select the framerate they want to show...I think something which could be cool would be allow the user to pause or rewind the running gif to investigate a cool or interesting movement pattern.
```{r}
# anim  = sample_play %>% ggplot() + geom_point(aes(x = x, y = y, color = team)) + theme(legend.position = "none") +
#         transition_time(frameId)  +
#         ease_aes('linear') +
#         NULL
# nFrames <- max(sample_play$frameId)
# anim_save('LowestOffsetAnimation.gif',
#           animate(anim, width = 720, height = 440,
#                   fps = 10, nframe = nFrames))

# spits out an animation gif
create_animation <- function(play_df, gif_filename) {
  anim <- play_df %>%
    # ggplot() + geom_point(aes(x = x, y = y, color = team)) + theme(legend.position = "none") +
    ggplot() +
    theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), panel.background = element_rect(fill = "#c2efff")) +
    xlim(0, 120) +
    ylim(0, 53.33) +

    ## NFL Field Specific Lines
    # add yard lines every 10 yds
    annotate("segment", x = seq(20, 105, by = 10), y = 0, xend = seq(20, 105, by = 10), yend = 53.33, col = "black") +

    # set mid-field and endzone lines
    annotate("segment", x = c(60), y = 0, xend = c(60), yend = 53.33, lwd = 1.3, col = "white") +
    annotate("segment", x = c(0, 10, 110, 120), y = 0, xend = c(0, 10, 110, 120), yend = 53.33, lwd = 1.1, col = "black") +

    # add lines at locations of hash marks
    annotate("segment", x = 10, y = (29.24667 + 30.24667) / 2, xend = 110, yend = (29.24667 + 30.24667) / 2, color = "white", lwd = 1.05) +
    annotate("segment", x = 10, y = (24.08333 + 23.08333) / 2, xend = 110, yend = (24.08333 + 23.08333) / 2, color = "white", lwd = 1.05) +

    # sideline
    annotate("segment", x = 0, y = c(0, 53.33), xend = 120, yend = c(0, 53.33), lwd = 1.3, col = "black") +
    # denote 50yd, 30yd, 10yd line with text
    annotate("text", label = 50, x = 60, y = 10, colour = "black", size = 5) +
    annotate("text", label = 30, x = 80, y = 6, colour = "black", size = 5) +
    annotate("text", label = 10, x = 100, y = 2, colour = "black", size = 5) +
    annotate("text", label = 30, x = 40, y = 6, colour = "black", size = 5) +
    annotate("text", label = 10, x = 20, y = 2, colour = "black", size = 5) +
    # above labels
    annotate("text", label = 50, x = 60, y = 44, colour = "black", size = 5) +
    annotate("text", label = 30, x = 80, y = 48, colour = "black", size = 5) +
    annotate("text", label = 10, x = 100, y = 51.5, colour = "black", size = 5) +
    annotate("text", label = 30, x = 40, y = 48, colour = "black", size = 5) +
    annotate("text", label = 10, x = 20, y = 51.5, colour = "black", size = 5) +

    # field goal posts
    annotate(geom = "segment", x = 120, xend = 120, y = 160 / 6 + 18.5 / 3 / 2, yend = 160 / 6 - 18.5 / 3 / 2, color = "yellow", lwd = 1.3) +
    annotate(geom = "segment", x = 0, xend = 0, y = 160 / 6 + 18.5 / 3 / 2, yend = 160 / 6 - 18.5 / 3 / 2, color = "yellow", lwd = 1.3) +

    ## add the points last so they sit ontop of the field markings
    scale_color_manual(values = c("#ff1212", "#8f6545", "#057d11")) +
    geom_point(aes(x = x, y = y, color = team, shape = is_football), size = 3) +
    transition_time(frameId) +
    ease_aes("linear") +
    NULL

  nFrames <- max(play_df$frameId)
  anim_save(
    paste(gif_filename, ".gif", sep = ""),
    animate(anim,
      width = 720, height = 440,
      fps = 7, nframe = nFrames
    )
  )
}
```

There will be a drop down where the user can select the special teams play type (kickoff, punt, field goal, extra point). We will create gifs of some of the best results

# Create Gifs 

## Extra Point
```{r, extra-point}
## extra point - not exciting...select one for each result (kick attempt good, blocked, kick attempt no good)
extra_point_outcomes <- plays %>%
  filter(specialTeamsPlayType == "Extra Point" & gameId > 2020000000) %>%
  group_by(specialTeamsResult) %>%
  slice(n = 1)

# add extra columns for gif processing
extra_point_outcomes$game_play_key <- paste(extra_point_outcomes$gameId, "_", extra_point_outcomes$playId, sep = "")
extra_point_outcomes$gif_name <- str_replace_all(paste(extra_point_outcomes$specialTeamsPlayType, "_", extra_point_outcomes$specialTeamsResult, sep = ""), fixed(" "), "")

# grab the location data for the specified plays
extra_point <- extra_point_outcomes %>% inner_join(tracking, by = c("gameId", "playId"))
extra_point <- extra_point %>%
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))

# end data augmentation

# create gif for each unique game/play key
for (i in unique(extra_point$game_play_key)) {
  play_df <- extra_point %>% filter(game_play_key == i)
  create_animation(play_df, unique(play_df$gif_name))
}
```

## Punt
```{r, punt}
punt_outcomes <- plays %>%
  filter(specialTeamsPlayType == "Punt" & gameId > 2020000000) %>%
  group_by(specialTeamsResult) %>%
  slice(n = 1)
# add extra columns for gif processing
punt_outcomes$game_play_key <- paste(punt_outcomes$gameId, "_", punt_outcomes$playId, sep = "")
punt_outcomes$gif_name <- str_replace_all(paste(punt_outcomes$specialTeamsPlayType, "_", punt_outcomes$specialTeamsResult, sep = ""), fixed(" "), "")

# grab the location data for the specified plays
punt <- punt_outcomes %>% inner_join(tracking, by = c("gameId", "playId"))
punt <- punt %>%
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))

# end data augmentation

# create gif for each unique game/play key
for (i in unique(punt$game_play_key)) {
  play_df <- punt %>% filter(game_play_key == i)
  create_animation(play_df, unique(play_df$gif_name))
}
```

## Kickoff
```{r}
kickoff_outcomes <- plays %>%
  filter(specialTeamsPlayType == "Kickoff" & gameId > 2020000000) %>%
  group_by(specialTeamsResult) %>%
  slice(n = 1)
# add extra columns for gif processing
kickoff_outcomes$game_play_key <- paste(kickoff_outcomes$gameId, "_", kickoff_outcomes$playId, sep = "")
kickoff_outcomes$gif_name <- str_replace_all(paste(kickoff_outcomes$specialTeamsPlayType, "_", kickoff_outcomes$specialTeamsResult, sep = ""), fixed(" "), "")

# grab the location data for the specified plays
kickoff <- kickoff_outcomes %>% inner_join(tracking, by = c("gameId", "playId"))
kickoff <- kickoff %>%
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))

# end data augmentation

# create gif for each unique game/play key
for (i in unique(kickoff$game_play_key)) {
  play_df <- kickoff %>% filter(game_play_key == i)
  create_animation(play_df, unique(play_df$gif_name))
}
```

## Field Goal
```{r}
field_goal_outcomes <- plays %>%
  filter(specialTeamsPlayType == "Field Goal" & gameId > 2020000000) %>%
  group_by(specialTeamsResult) %>%
  slice(n = 1)
# add extra columns for gif processing
field_goal_outcomes$game_play_key <- paste(field_goal_outcomes$gameId, "_", field_goal_outcomes$playId, sep = "")
field_goal_outcomes$gif_name <- str_replace_all(paste(field_goal_outcomes$specialTeamsPlayType, "_", field_goal_outcomes$specialTeamsResult, sep = ""), fixed(" "), "")

# grab the location data for the specified plays
field_goal <- field_goal_outcomes %>% inner_join(tracking, by = c("gameId", "playId"))
field_goal <- field_goal %>%
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))

# end data augmentation

# create gif for each unique game/play key
for (i in unique(field_goal$game_play_key)) {
  play_df <- field_goal %>% filter(game_play_key == i)
  create_animation(play_df, unique(play_df$gif_name))
}
```

