---
title: "AnimateGraph"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AnimateGraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(popo)
library(dplyr)
library(stringr)
library(ggplot2)
library(gganimate)
library(tidyverse)
library(cowplot)
library(ggridges)
library(repr)

#turning off warnings
options(warn=-1)

#setting plot width and height
options(repr.plot.width=15, repr.plot.height = 10)
```

Function to plot the field ('https://raw.githubusercontent.com/mlfurman3/gg_field/main/gg_field.R')
```{r}
# field function to reduce redundancy of plotting code on annotations
create_field <- function(){
  
}
```



Read in Raw Data
```{r}
## run code in console
plays = read.csv(file = paste(getwd(),"/data/plays.csv",sep=''), header = TRUE)
tracking = read.csv(file = paste(getwd(),"/data/tracking2020.csv",sep=''), header = TRUE)
##
```

Grab one play for plot testing:
```{r}
sample_gameId = "2020092012"
sample_playId = "1357"
sample_play = tracking %>% filter(gameId == sample_gameId & playId == sample_playId)

# add a group (is football or not) 
sample_play = sample_play %>% 
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))

## clean up sample plot for animation
sample_play %>% filter(frameId == 1) %>% 
  ggplot() + theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(), panel.background = element_rect(fill = "#c2efff")) +
  xlim(0,120) + ylim(0, 53.33) +
  
  ## NFL Field Specific Lines
  # add yard lines every 10 yds
  annotate("segment", x=seq(20,105,by=10), y=0, xend=seq(20,105,by=10), yend=53.33, col="black") +
  
  # set mid-field and endzone lines
  annotate("segment",x=c(60), y=0, xend=c(60), yend=53.33, lwd=1.3, col="white") +
  
  annotate("segment",x=c(0,10,110,120), y=0, xend=c(0,10,110,120), yend=53.33, lwd=1.1, col="black") +
  
  # add lines at locations of hash marks
  annotate("segment", x = 10, y=(29.24667 + 30.24667)/2, xend=110, yend=(29.24667 + 30.24667)/2, color="white", lwd=1.05) +
  annotate("segment", x=10, y=(24.08333 + 23.08333)/2, xend=110, yend=(24.08333 + 23.08333)/2, color="white", lwd=1.05) +
  
  # sideline
  annotate("segment",x=0, y=c(0, 53.33), xend=120, yend=c(0, 53.33), lwd=1.3, col="black") +
  # denote 50yd, 30yd, 10yd line with text
  annotate("text", label=50, x=60, y=10, colour="black", size=5) + 
  annotate("text", label=30, x=80, y=6, colour="black", size=5) + 
  annotate("text", label=10, x=100, y=2, colour="black", size=5) + 
  annotate("text",label=30,x=40, y=6, colour="black", size=5) + 
  annotate("text",label=10,x=20, y=2, colour="black", size=5) + 
  # above labels
  annotate("text",label=50,x=60, y=44, colour="black", size=5) + 
  annotate("text",label=30,x=80, y=48, colour="black", size=5) + 
  annotate("text",label=10,x=100, y=51.5, colour="black", size=5) + 
  annotate("text",label=30,x=40, y=48, colour="black", size=5) + 
  annotate("text",label=10,x=20, y=51.5, colour="black", size=5) + 
  
  # field goal posts
  annotate(geom = 'segment', x = 120, xend = 120, y = 160/6 + 18.5/3/2, yend = 160/6 - 18.5/3/2, color = "yellow", lwd = 1.3) +
  annotate(geom = 'segment', x = 0, xend = 0, y = 160/6 + 18.5/3/2, yend = 160/6 - 18.5/3/2, color = "yellow", lwd = 1.3) +

  ## add the points last so they sit ontop of the field markings
  scale_color_manual(values = c("#ff1212", "#8f6545", "#057d11")) +
  geom_point(aes(x = x, y = y, color = team, shape = is_football), size = 3) 
  

# uses user function to plot the field
# sample_play %>% filter(frameId == 1) %>% ggplot() + gg_field() + geom_point(aes(x = x, y = y, color = team))
```

Create Test Animation...using a series of .png files use R libary to string together an animation.
We will serve this on the webpage. Could also have option for user to select the framerate they want to show...I think something which could be cool would be allow the user to pause or rewind the running gif to investigate a cool or interesting movement pattern.
```{r}
anim  = sample_play %>% ggplot() + geom_point(aes(x = x, y = y, color = team)) + theme(legend.position = "none") +
        transition_time(frameId)  +
        ease_aes('linear') + 
        NULL
nFrames <- max(sample_play$frameId)
anim_save('LowestOffsetAnimation.gif',
          animate(anim, width = 720, height = 440,
                  fps = 10, nframe = nFrames))

# spits out an animation gif
create_animation = function(play_df, gif_filename){
  
  anim  = play_df %>% 
    # ggplot() + geom_point(aes(x = x, y = y, color = team)) + theme(legend.position = "none") +
    ggplot() + theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(), panel.background = element_rect(fill = "#c2efff")) +
  xlim(0,120) + ylim(0, 53.33) +
  
  ## NFL Field Specific Lines
  # add yard lines every 10 yds
  annotate("segment", x=seq(20,105,by=10), y=0, xend=seq(20,105,by=10), yend=53.33, col="black") +
  
  # set mid-field and endzone lines
  annotate("segment",x=c(60), y=0, xend=c(60), yend=53.33, lwd=1.3, col="white") +
  
  annotate("segment",x=c(0,10,110,120), y=0, xend=c(0,10,110,120), yend=53.33, lwd=1.1, col="black") +
  
  # add lines at locations of hash marks
  annotate("segment", x = 10, y=(29.24667 + 30.24667)/2, xend=110, yend=(29.24667 + 30.24667)/2, color="white", lwd=1.05) +
  annotate("segment", x=10, y=(24.08333 + 23.08333)/2, xend=110, yend=(24.08333 + 23.08333)/2, color="white", lwd=1.05) +
  
  # sideline
  annotate("segment",x=0, y=c(0, 53.33), xend=120, yend=c(0, 53.33), lwd=1.3, col="black") +
  # denote 50yd, 30yd, 10yd line with text
  annotate("text", label=50, x=60, y=10, colour="black", size=5) + 
  annotate("text", label=30, x=80, y=6, colour="black", size=5) + 
  annotate("text", label=10, x=100, y=2, colour="black", size=5) + 
  annotate("text",label=30,x=40, y=6, colour="black", size=5) + 
  annotate("text",label=10,x=20, y=2, colour="black", size=5) + 
  # above labels
  annotate("text",label=50,x=60, y=44, colour="black", size=5) + 
  annotate("text",label=30,x=80, y=48, colour="black", size=5) + 
  annotate("text",label=10,x=100, y=51.5, colour="black", size=5) + 
  annotate("text",label=30,x=40, y=48, colour="black", size=5) + 
  annotate("text",label=10,x=20, y=51.5, colour="black", size=5) + 
  
  # field goal posts
  annotate(geom = 'segment', x = 120, xend = 120, y = 160/6 + 18.5/3/2, yend = 160/6 - 18.5/3/2, color = "yellow", lwd = 1.3) +
  annotate(geom = 'segment', x = 0, xend = 0, y = 160/6 + 18.5/3/2, yend = 160/6 - 18.5/3/2, color = "yellow", lwd = 1.3) +

  ## add the points last so they sit ontop of the field markings
  scale_color_manual(values = c("#ff1212", "#8f6545", "#057d11")) +
  geom_point(aes(x = x, y = y, color = team, shape = is_football), size = 3) +
        transition_time(frameId)  +
        ease_aes('linear') + 
        NULL
  
  nFrames <- max(play_df$frameId)
  anim_save(paste(gif_filename, '.gif', sep=''),
          animate(anim, width = 720, height = 440,
                  fps = 7, nframe = nFrames))
}
```

There will be a drop down where the user can select the special teams play type (kickoff, punt, field goal, extra point). We will create gifs of some of the best results
```{r}
## extra point - not exciting...select one for each result (kick attempt good, blocked, kick attempt no good)
extra_point_outcomes = plays %>% filter(specialTeamsPlayType == "Extra Point" & gameId > 2020000000) %>% group_by(specialTeamsResult) %>% slice(n = 1)

# add extra columns for gif processing
extra_point_outcomes$game_play_key = paste(extra_point_outcomes$gameId, "_", extra_point_outcomes$playId, sep='')
extra_point_outcomes$gif_name = str_replace_all(paste(extra_point_outcomes$specialTeamsPlayType, "_", extra_point_outcomes$specialTeamsResult, sep=''), fixed(" "), "")


# grab the location data for the specified plays
extra_point = extra_point_outcomes %>% inner_join(tracking, by = c("gameId", "playId"))
extra_point = extra_point %>% 
  mutate(is_football = case_when(displayName == "football" ~ TRUE, TRUE ~ FALSE))
# end data augmentation

# create gif for each unique game/play key
for(i in unique(extra_point$game_play_key)){
  play_df = extra_point %>% filter(game_play_key == i)
  create_animation(play_df, unique(play_df$gif_name))
}
```



